---
title: "Merged Final Econ Project"
output: pdf_document
date: "2023-12-11"
---

## Load libraries
```{r}
library(readr)
library(glmnet)
library(corrplot)
library(Metrics)
library(dplyr)
library(ggplot2)
library(scales)
library(lubridate)
```


## Importing the dataset

```{r}
df_store <- read_csv("~/University/E401 Machine Learning For Economic Data/Final Project/Walmart Sales Forecast/stores.csv")
df_features = read_csv("~/University/E401 Machine Learning For Economic Data/Final Project/Walmart Sales Forecast/features.csv") 
df_train = read_csv("~/University/E401 Machine Learning For Economic Data/Final Project/Walmart Sales Forecast/train.csv")
```

## Preprocessing of the dataset

```{r}
# Display the first three rows of the data frame 
head(df_store, 3)
head(df_features, 3)

#merging df_train and df_features dataset and df_store
df <- inner_join(inner_join(df_train, df_features, by = c("Store", "Date")), df_store, by = "Store")

# Print the first 5 rows of the resulting data frame
head(df, 5)
# removing duplicate column since IsHoliday_y and IsHoliday_x are the same
df$IsHoliday.y <- NULL
#renaming Isholiday.x as isHoliday
names(df)[names(df) == "IsHoliday.x"] <- "IsHoliday"
# Print the first 5 rows of the resulting data frame
head(df, 5)

# check for non-zero and zero values for weekly sales
filtered_df <- subset(df, Weekly_Sales <= 0)
filtered_df

# total rows in dataframe = 421570
# total rows with missing or zero or negative values = 1348
# perventage of rows with missing values= 1348/ 421570= 0.31%

# therefore, removing them 
df <- subset(df, Weekly_Sales > 0)
df

colnames(df)[colnames(df) == "IsHoliday_x"] <- "IsHoliday"
df
```
Data Description
```{r}
summary(df)
```


## EXPLORATORY DATA ANALYSIS

Holiday vs. Weekly Sales
```{r}

# Create a bar plot
ggplot(df, aes(x = IsHoliday, y = Weekly_Sales, fill = IsHoliday)) +
  geom_bar(stat = "identity") +
  labs(x = "Holiday", y = "Weekly Sales") +
  ggtitle("Holiday vs. Weekly Sales")
```

Print the dates of holidays
```{r}

# Create a subset for rows where 'IsHoliday' is true
df_holiday <- subset(df, IsHoliday == TRUE)

# Get unique dates from the subset
unique_dates <- unique(df_holiday$Date)

# Print unique dates
print(unique_dates)
```
Super Bowl, Labor Day, Thanksgiving, Christmas are the holidays tht are 
present in the dataset.

Create different rows in the dataset for the unique holidays
```{r}

#create different rows in the dataset for the unique holidays
df$Super_Bowl <- ifelse(df$Date %in% as.Date(c('2010-02-12', '2011-02-11', '2012-02-10')), TRUE, FALSE)
df$Thanksgiving <- ifelse(df$Date %in% as.Date(c('2010-11-26', '2011-11-25')), TRUE, FALSE)
df$Labor_Day <- ifelse(df$Date %in% as.Date(c('2010-09-10', '2011-09-09', '2012-09-07')), TRUE, FALSE)
df$Christmas <- ifelse(df$Date %in% as.Date(c('2010-12-31', '2011-12-30')), TRUE, FALSE)

```

THANKSGIVING AVERAGE SALES VS AVERAGE WEEKLY SALES
```{r}

# Calculate the average weekly sales for both TRUE and FALSE values
average_sales <- tapply(df$Weekly_Sales, df$Thanksgiving, mean)

# Create a data frame for plotting
plot_data <- data.frame(
  Thanksgiving = factor(names(average_sales)),
  Average_Weekly_Sales = average_sales
)

# Create a bar plot
ggplot(plot_data, aes(x = Thanksgiving, y = Average_Weekly_Sales, fill = Thanksgiving)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Thanksgiving", y = "Average Weekly Sales") +
  scale_fill_manual(values = c("FALSE" = "blue", "TRUE" = "red"), name = "Thanksgiving") +
  ggtitle("Thanksgiving vs. Average Weekly Sales")
```


LABOUR DAY AVERAGE SALES VS AVERAGE WEEKLY SALES
```{r}

# Calculate the average weekly sales for both TRUE and FALSE values for Labor Day
average_sales_labor_day <- tapply(df$Weekly_Sales, df$Labor_Day, mean)

# Create a data frame for plotting for Labor Day
plot_data_labor_day <- data.frame(
  Holiday = factor(names(average_sales_labor_day)),
  Average_Weekly_Sales = average_sales_labor_day
)

# Create a bar plot for Labor Day
ggplot(plot_data_labor_day, aes(x = Holiday, y = Average_Weekly_Sales, fill = Holiday)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Labor Day", y = "Average Weekly Sales") +
  scale_fill_manual(values = c("FALSE" = "blue", "TRUE" = "red"), name = "Labor Day") +
  ggtitle("Labor Day vs. Average Weekly Sales")

```

Christmas vs. Average Weekly Sales
```{r}

# Calculate the average weekly sales for both TRUE and FALSE values for Christmas
average_sales_christmas <- tapply(df$Weekly_Sales, df$Christmas, mean)

# Create a data frame for plotting for Christmas
plot_data_christmas <- data.frame(
  Holiday = factor(names(average_sales_christmas)),
  Average_Weekly_Sales = average_sales_christmas
)

# Create a bar plot for Christmas
ggplot(plot_data_christmas, aes(x = Holiday, y = Average_Weekly_Sales, fill = Holiday)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Christmas", y = "Average Weekly Sales") +
  scale_fill_manual(values = c("FALSE" = "blue", "TRUE" = "red"), name = "Christmas") +
  ggtitle("Christmas vs. Average Weekly Sales")


```


Pie chart withthe average weekly sales by store type
```{r}

# Calculate the average weekly sales by store type
average_sales_by_store <- aggregate(Weekly_Sales ~ Type, data = df, FUN = mean)

# Store types
store_types <- average_sales_by_store$Type

# Average weekly sales data
average_sales_data <- average_sales_by_store$Weekly_Sales

# Plotting a pie chart with percentage labels
pie(average_sales_data, labels = paste0(store_types, "\n", percent(average_sales_data / sum(average_sales_data))),
    col = c("blue", "orange", "green"), cex = 1.2, border = "white",
    main = "Average Weekly Sales by Store Type", cex.lab = 1.5)

```

Boxplot of Size of Store by Type of Store
```{r}

# Create a boxplot
ggplot(df_store, aes(x = Type, y = Size)) +
  geom_boxplot() +
  labs(title = "Boxplot of Size by Type", x = "Type", y = "Size")
```

Scatter plot for 'month' and 'year' against average weekly sales
```{r}

df$Date <- as.Date(df$Date)

# Create 'week', 'month', and 'year' columns
df$week <- week(df$Date)
df$month <- month(df$Date)
df$year <- year(df$Date)


# Calculate average weekly sales for each combination of 'month' and 'year'
df_avg <- df %>%
  group_by(month, year) %>%
  summarise(Avg_Weekly_Sales = mean(Weekly_Sales, na.rm = TRUE))

# Create a scatter plot for 'month' and 'year' against average weekly sales
ggplot(df_avg, aes(x = month, y = Avg_Weekly_Sales, color = factor(year))) +
  geom_line() +
  labs(title = "Average Weekly Sales by Month and Year", x = "Month", y = "Average Weekly Sales") +
  scale_color_manual(values = c("blue", "red", "green")) +  # Set your desired color for each year
  theme_minimal()
```

Average Weekly Sales Against Unemployment
```{r}

average_sales <- df %>%
  group_by(Unemployment) %>%
  summarise(Avg_Weekly_Sales = mean(Weekly_Sales, na.rm = TRUE))

# Create a line plot with 'Unemployment' on the x-axis and average 'Weekly_Sales' on the y-axis
ggplot(average_sales, aes(x = Unemployment, y = Avg_Weekly_Sales)) +
  geom_line(color = "blue") +
  labs(title = "Average Weekly Sales Against Unemployment", x = "Unemployment", y = "Average Weekly Sales") +
  theme_minimal()

```


Average Weekly Sales Against Unemployment
```{r}

average_sales <- df %>%
  group_by(CPI) %>%
  summarise(Avg_Weekly_Sales = mean(Weekly_Sales, na.rm = TRUE))

ggplot(average_sales, aes(x = CPI, y = Avg_Weekly_Sales)) +
  geom_line(color = "blue") +
  labs(title = "Average Weekly Sales Against CPI", x = "CPI", y = "Average Weekly Sales") +
  theme_minimal()
```


Average Weekly Sales Against Temperature
```{r}

average_sales <- df %>%
  group_by(Temperature) %>%
  summarise(Avg_Weekly_Sales = mean(Weekly_Sales, na.rm = TRUE))

ggplot(average_sales, aes(x = Temperature, y = Avg_Weekly_Sales)) +
  geom_line(color = "blue") +
  labs(title = "Average Weekly Sales Against Temperature", x = "Temperature", y = "Average Weekly Sales") +
  theme_minimal()

```

```{r}
average_sales <- df %>%
  group_by(Fuel_Price) %>%
  summarise(Avg_Weekly_Sales = mean(Weekly_Sales, na.rm = TRUE))

ggplot(average_sales, aes(x = Fuel_Price, y = Avg_Weekly_Sales)) +
  geom_line(color = "blue") +
  labs(title = "Average Weekly Sales Against Unemployment", x = "Unemployment", y = "Average Weekly Sales") +
  theme_minimal()

```

```{r}
average_sales <- df %>%
  group_by(month) %>%
  summarise(Avg_Weekly_Sales = mean(Weekly_Sales, na.rm = TRUE))
# Create a bar plot with 'month' on the x-axis and 'Weekly_Sales' on the y-axis
ggplot(average_sales, aes(x = month, y = Avg_Weekly_Sales)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "Monthly Sales", x = "Month", y = "Weekly Sales") +
  theme_minimal()


```

Let's look at the correlation plot
```{r}
data <- read_csv("~/University/E401 Machine Learning For Economic Data/Final Project/Walmart Sales Forecast/data5.csv")
corr = cor(data[, c(1:9)])
corrplot(corr, method = "color", cl.pos = 'n', rect.col = "black",  tl.col = "black", addCoef.col = "black", number.digits = 2, number.cex = 0.60, tl.cex = 0.7, cl.cex = 1, col = colorRampPalette(c("darkred","white","lightblue"))(100))
```

## Econometric Methods

Read the pre-processed dataset
```{r}
train <- read_csv("~/University/E401 Machine Learning For Economic Data/Final Project/Walmart Sales Forecast/trainingSet.csv")
test <- read_csv("~/University/E401 Machine Learning For Economic Data/Final Project/Walmart Sales Forecast/testSet.csv")
```

Prepare the train and test dataset
```{r}
y_train <- train$Weekly_Sales
X_train <- subset(train, select = -Weekly_Sales)
X_train <- as.matrix(X_train)
y_train <- as.matrix(y_train)
y_test <- test$Weekly_Sales
X_test <- subset(test, select = -Weekly_Sales)
```

Perform Linear Regression
```{r}
fit_lm <- lm(Weekly_Sales ~ ., data = train)
summary(fit_lm)
```

Check performance against test set
```{r}
prediction = predict(fit_lm, newdata = X_test)
residuals <- y_test - prediction
squared_residuals <- residuals^2
mse <- mean(squared_residuals)
rmse_lm <- sqrt(mse)
rmse_lm
```

```{r}
mae_lm <- mae(y_test, prediction)
mae_lm
```

Let's perform the same but with LASSO
```{r}
lasso_model <- cv.glmnet(X_train, y_train, alpha = 1)
coef(lasso_model)
```

Prepare the test dataset
```{r}
X_test <- as.matrix(X_test)
y_test <- as.matrix(y_test)
```

```{r}
prediction <- predict(lasso_model, newx = X_test)
mae_lasso <- mae(y_test, prediction)
mae_lasso
```

```{r}
plot(lasso_model)
```

One interesting analysis we can perform is see if removing the features that we thought were promising from EDA will impact the 
performance of our model.
To begin with, we can try removing the holidays information from our training dataset 
```{r}
X_train_rem_hol <- subset(X_train, select = -EventsNo_Holiday)
X_train_rem_hol <- as.matrix(X_train_rem_hol)
fit2_rem_hol <- lm(y_train ~ X_train_rem_hol)
summary(fit2_rem_hol)
```
Looks like that one did not make too much of a difference

We remember seeing that the weekly sales depended on which month it was too
```{r}
X_train_rem_months <- X_train[, !(colnames(X_train) %in% c("month2", "month3", "month4", "month5", "month6", "month7", "month8", "month9", "month10", "month11", "month12"))]
X_train_rem_months <- as.matrix(X_train_rem_months)
fit3 <- lm(y_train ~ X_train_rem_months)
summary(fit3)
```
We see that the adjusted R-squared value has dropped significantly. So our EDA has proved beneficial in keeping the month variable even though correlation did not show the same. 

```{r}
X_train_rem_fuel <- X_train[, !(colnames(X_train) %in% c("Fuel_Price"))]
X_train_rem_fuel <- as.matrix(X_train_rem_fuel)
fit4 <- lm(y_train ~ X_train_rem_fuel)
summary(fit4)
```

```{r}
X_train_rem_emp <- X_train[, !(colnames(X_train) %in% c("Unemployment"))]
X_train_rem_emp <- as.matrix(X_train_rem_emp)
fit5 <- lm(y_train ~ X_train_rem_emp)
summary(fit5)
```
Random Forests
```{r}
library(randomForest)

# Fit random forest model using the training data
rf_model <- randomForest(X_train, y_train, ntree = 500, mtry = 4)

# View model summary
print(rf_model)
```

RMSE
```{r}
# Predict using the trained model on the test data
predictions <- predict(rf_model, X_test)

# Calculate RMSE (if needed)
# Assuming 'y_test' contains the actual test set values
rmse_rf <- sqrt(mean((predictions - y_test)^2))
rmse_rf
```

```{r}
plot(rf_model)
```



